---
title: "Welcome to KIN 610!"
subtitle: "KIN 610 - Spring 2023"
author: "Dr. Ovande Furtado Jr"
footer:  "[https://drfurtado.github.io/kin610/](https://drfurtado.github.io/kin610/)"
logo: "images/logo.png"
format: 
  revealjs: 
    theme: slides.scss
    multiplex: true
    transition: fade
    slide-number: true
    width: 16:9
editor: visual
execute:
  freeze: auto
---

```{r include = FALSE}
library(countdown)
```

## Meet each other!

In breakout rooms:

-   Name, year, major, hometown
-   What did you do over the winter break?
-   What do you hope to get out of this course?
-   Anything else you want to share/ask?

```{r}
countdown(minutes = 5)
```

## **What to expect in class** {.scrollable .incremental}

-   ::: incremental
    -   Try to arrive 10 minutes before class begins to set up the computer, open lecture notes, etc.

    -   I will go over any announcements and begin the lecture shortly after

    -   A PDF version of the lecture notes will be provide to students, which can used for note taking purposes - electronically[^1] or hand writing[^2]

    -   When covering jamovi, students are expected to follow along and turn in the generated output at the end of class - will count toward participation points
    :::

[^1]: Use Acrobat Reader for this purpose.

[^2]: The computer lab has a printer.

## Wrap up

Are there any questions {{< bi question-diamond-fill >}}
