---
title: "Statistical Tables"
---

::: panel-tabset
## Tables

The field of statistics relies heavily on mathematical formulas and calculations to make sense of data. Statistical tables, such as the z-table, t-table, f-table, and Pearson table, are essential tools for statisticians and researchers. These tables provide critical values for different probability distributions, enabling users to perform statistical tests, calculate confidence intervals, and make informed decisions based on data analysis.

## z-table

The table shows the probabilities associated with various z-scores. Each row represents a different z-score, while the columns show the z-score value and the probability (rounded to two decimal places) of getting a z-score less than or equal to that value.

For example, if we look at the row where z is equal to 1.96, we see that the probability of getting a z-score less than or equal to 1.96 is 0.025 (1 - 0.9750). This means that if we have a normally distributed variable, and we want to find the probability of getting a z-score less than or equal to 1.96, we can look up that value in the table and find the corresponding probability. In this case, the probability is 0.9750, then subtract from 1. The result is the probability value.

Conversely, if we have a probability and we want to find the z-score associated with it, we can use the table by looking for the probability in the second column and then reading the corresponding z-score from the first column. For example, if we have a probability of 0.9985, we can look for that value in the second column and find the corresponding z-score of 2.81 in the first column. Overall, this table can be a useful tool for working with the standard normal distribution, particularly when calculating probabilities or finding critical values for hypothesis tests.

```{r, echo=FALSE}
# Set up a vector of z values
z <- seq(0.00, 3.40, 0.01)

# Calculate the probabilities for each z value using pnorm()
probabilities <- round(pnorm(z), 4)

# Combine the z values and probabilities into a data frame
table <- data.frame(z, probabilities)

# Print the table
print(table)


```

## correlation

```{r, echo=FALSE}
# set up the degrees of freedom (df) range
df <- 1:99

# set up the alpha values
alpha <- c(0.10, 0.05, 0.025, 0.01)

# set up an empty matrix to store the critical values
crit_vals <- matrix(NA, nrow = length(df), ncol = length(alpha)+1)

# fill in the first column of the matrix with degrees of freedom
crit_vals[,1] <- df

# calculate the critical values and store them in the matrix
for (i in 1:length(df)) {
  for (j in 1:length(alpha)) {
    crit_vals[i,j+1] <- round(qt(1-alpha[j]/2, df[i]), 3)
  }
}

# set the column names
colnames(crit_vals) <- c("df", paste0("alpha=", alpha))

# print the table
crit_vals

```

:::

